\chapter{Introduction}\label{ch:introduction}

TaskForce is a distributed job scheduling system designed to process complex heterogeneous tasks across a polyglot cluster.
The system demonstrates the practical application of fundamental distributed systems concepts,
including consensus algorithms, message-oriented middleware, fault tolerance, and distributed coordination, fulfilling all requirements of the Distributed Systems course.

The project implements a complete job scheduling pipeline featuring:
\begin{itemize}
    \item \textbf{Erlang/OTP Orchestrator}\cite{erlang}: A robust coordinator utilizing the Raft consensus algorithm for leader election and CRDTs for state management.
    \item \textbf{RabbitMQ Cluster}\cite{rabbitmq}: Message-Oriented Middleware providing reliable, priority-based asynchronous queuing.
    \item \textbf{Java Spring Boot API Gateway}\cite{springboot}: A web-facing entry point with REST endpoints and real-time Server-Sent Events (SSE) streaming.
    \item \textbf{Polyglot Worker Farm}: Heterogeneous execution nodes supporting Python and Java implementations.
    \item \textbf{Multi-node Deployment}: Distributed across Ubuntu VMs (Adrien0, Adrien1) with automatic failover.
\end{itemize}

The system successfully handles concurrent job submissions, automatically recovers from node failures, maintains reliable execution semantics, and scales horizontally.

\section{Background}\label{sec:background}

Distributed systems have become the backbone of modern computing infrastructure.
From cloud computing platforms to microservices architectures, the ability to coordinate work across multiple nodes while maintaining consistency and availability is fundamental.
Job scheduling-the allocation of computational tasks to worker nodes-represents a classic distributed systems problem that encompasses core challenges: coordination, communication, fault tolerance, and load balancing.

\section{Problem Statement}\label{sec:problem-statement}

Modern distributed applications require reliable job processing systems that can:
\begin{itemize}
    \item Handle concurrent task submissions asynchronously.
    \item Guarantee fault tolerance and state consistency across nodes.
    \item Scale horizontally as workload increases.
    \item Support heterogeneous (polyglot) execution environments.
\end{itemize}
This project aims to build a complete distributed job scheduler from first principles, demonstrating how fundamental distributed algorithms and middleware technologies address these challenges.

\section{Project Scope}\label{sec:project-scope}

The scope of this project includes the design and implementation of a distributed scheduler, Raft consensus for leader election, AMQP integration, a polyglot worker framework, and comprehensive API design.


\vspace{1em}
\noindent \textbf{Project Repository:} \\
The complete source code for TaskForce, including the Erlang orchestrator, Java API Gateway, polyglot worker implementations, and deployment scripts, is publicly available on GitHub.
It can be accessed at: \\
\url{https://github.com/adrienKoumgangT/distriqueue}

